An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2021-08-25 latex_files/ch2/2_4_ROMC.tex
changed notation to be inline with ROMC paper

\begin{itemize}
\item The whole acceptable region $\mathcal{S}_i$
  shrinks to a single point $\thetab_i^*$/. This simplification adds
  significant apprimation error when then the area $\mathcal{S}_i$ is relatively big.
\item The weight $w_i$ is computed based only at the curvature at the
  point $\thetab_i^*$. Using only local information for estimating the weight can also be misleading. For example, if $g_i$ is almost flat at $\thetab_i^*$, i.e. $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0, then \Rightarrow w_i
  \rightarrow \infty$ dominating the posterior.
\item There is no way to solve the optimisation problem
  $\thetab_i^* = \text{argmin}_{\thetab} \: g_i(\thetab)$ when $g_i$
  is not differentiable.
\end{itemize}

@@

ROMC resolves the failure-modes of OMC mainly by approaching the
acceptance regions $\mathcal{S}_i$ in a more sophisticated
way. Instead of collapsing the acceptance regions into single points,
it approximates them with a bounding box region. A uniform distribution
defined on the bounding box area, is then used as the \emph{proposal
  distribution} for importance sampling. If $q_i$ is the uniform
distribution defined on the $i-th$ bounding box, weighted sampling is
performed by drawing samples from $q_i$ and assigning the weights as
in~\eqref{eq:sampling}.

@@

For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$,
we search for a point $\thetab^*$ such that
$d(M_d(\thetab^*,\vb_i), \data) < \epsilon$. For notation convenience,
we name the function $d(M_d(\thetab^*,\vb_i)$ as $d_i(\thetab)$.  Obtaining
this point involves solving the following optimisation problem:

\begin{subequations}
\begin{alignat}{2}
  &\!\min_{\thetab}        &\qquad& d_i(\thetab) = d(M_d(\thetab^*,\vb_i), \data)\label{eq:optProb}\\
  &\text{subject to} & & d_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
The optimisation problem can be treated as unconstrained, accepting
the optimal point $\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)$
only if $d_i(\thetab_i^*) < \epsilon$. The nature of the generative
model $M_r(\thetab)$, specifies the properties of the objective
function $d_i$. If $d_i$ is differentiable any gradient-based
iterative algorithm can be used for~\ref{eq:optProb}. The gradients
$\nabla_{\thetab} g_i$ can be either provided in closed form or
approximated by finite differences. Alternatively, if $d_i$ is not
differentiable, the Bayesian Optimisation approach provides an
alternative choice~\cite{Shahriari2016}. In this case, apart from
obtaining an optimal $\thetab_i^* $, a surrogate model $\hat{d}_i$ of
the distance $d_i$ is also fitted; the approximate model can then
substitute the actual model in the following steps.

@@

After obtaining $\thetab_i^*$, we approximate the acceptance region
$\mathcal{S}_i$ by constructing a bounding box $\mathcal{\hat{S}}_i$
around it. The bounding must the part of the acceptance region
$\mathcal{S}_i$ around $\thetab_i^*$, formally
$\{ \thetab : d_i(\thetab) < \epsilon$,
$d(\thetab, \thetab_i^*) < M \}$. The second condition
$d(\thetab, \thetab_i^*) < M$ describes that if
$\mathcal{S}_i := \{ \thetab : d_i(\thetab) < \epsilon \} $ contains a
number of disjoint sets of $\thetab$ that they fulfill
$d_i(\thetab) < \epsilon$, we want the bounding box to match only the
set that contains $\thetab_i^*$. We seek for a bounding box that is as
tight as possible to the local acceptance region for ensuring high
acceptance rate in the importance sampling.

Bounding boxes are built in two steps. Firstly, the search directions
$\mathbf{v}_d$ are computed as the eigenvectors of the curvature at
$\thetab_i^*$ and then the limits are obtained with a line-search
method looking for the smallest
$\kappa: d_i(\thetab_i^* + \kappa \vb_d) \geq
\epsilon$\footnote{$-\kappa$ is used as well for the opposite
  direction along the search line}. The
Algorithm~\ref{alg:region_construction} describes the method
in-depth. After limits are found along all directions the uniform
distribution $q_i$ can be defined and is used as the proposal region
for importance sampling.

@@

After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
are no longer interested in $g_i(\thetab)$ outside
$\mathcal{\hat{S}}_i$. In the inference steps (e.g. sampling,
evaluating the posterior), we evaluate $g_i$ for checking if a sample
from the bounding box belongs to the acceptance region. Instead of
running the deterministic simulator $M_d(\thetab_{ij}, \vb_i)$ for
evaluating $g_i$, ROMC proposes fitting a \emph{light} surrogate model
$\hat{g}_i$ for approximating $g_i$ around $\theta_i^*$.

ROMC proposes fitting a simple quadratic model\footnote{Any ML
  regression model may be chosen as local surrogate. The choice should
  consider the properties of the local region (i.e. size,
  smoothness). } by sampling $N$ points from the bounding box and the
labels $Y: \mathbb{R}^{N}$ are computed by evaluating $g_i$. This
additional step adds an extra effort in the training part, but
promises a significant speed-up at the inference phase. Such imbalance
is usual in the context of ML where is usual to be quite generous with
the execution time at the training phase but quite eager at the
inference phase. Fitting a local surrogate model aligns with this
requirement.

>>>>

\begin{itemize}
\item The whole acceptable region $\mathcal{S}_i$
  shrinks to a single point $\thetab_i^*$. This simplification adds
  significant apprimation error when then the area $\mathcal{S}_i$ is relatively big.
\item The weight $w_i$ is computed based only at the curvature at the
  point $\thetab_i^*$. Using only local information for estimating the weight can also be misleading. For example, if $g_i$ is almost flat at $\thetab_i^*$, i.e. $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0, then \Rightarrow w_i
  \rightarrow \infty$ dominating the posterior.
\item There is no way to solve the optimisation problem
  $\thetab_i^* = \text{argmin}_{\thetab} \: g_i(\thetab)$ when $g_i$
  is not differentiable.
\end{itemize}

@@

ROMC resolves the failure-modes of OMC mainly by approaching the
acceptance regions $\mathcal{S}_i$ in a more sophisticated way.
Instead of collapsing the regions into single points, it approximates
them with a bounding box. Afterwards, a uniform distribution defined
on the bounding box area, is used as the \emph{proposal distribution}
for importance sampling. Defining as $q_i$ the uniform distribution
defined on the $i-th$ bounding box, weighted sampling is performed by
drawing samples from $q_i$ and assigning the weights as
in~\eqref{eq:sampling}.

@@

For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$,
we search for a point $\thetab^*_i$ such that
$d(M_d(\thetab^*_i,\vb_i), \data) < \epsilon$. For notation
convenience, we name the distance function $d(M_d(\thetab,\vb_i), \data)$ as
$d_i(\thetab)$.  Obtaining this point involves solving the following
optimisation problem:

\begin{subequations}
\begin{alignat}{2}
  &\!\min_{\thetab}        &\qquad& d_i(\thetab) = d(M_d(\thetab^*,\vb_i), \data)\label{eq:optProb}\\
  &\text{subject to} & & d_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
The optimisation problem can be treated as unconstrained, accepting
the optimal point $\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)$
only if $d_i(\thetab_i^*) < \epsilon$. The nature of the generative
model $M_r(\thetab)$, specifies the properties of the objective
function $d_i$. If $d_i$ is differentiable any gradient-based
iterative algorithm can be used for~\ref{eq:optProb}. The gradients
$\nabla_{\thetab} d_i$ can be either provided in closed form or
approximated by finite differences. Alternatively, if $d_i$ is not
differentiable, the Bayesian Optimisation
approach~\cite{Shahriari2016} provides an alternative choice. In this
case, apart from obtaining an optimal $\thetab_i^* $, a surrogate
model $\hat{d}_i$ of the distance $d_i$ is also fitted; the
approximate model can then substitute the actual model in the
following steps.

@@

After obtaining $\thetab_i^*$, we initially approximate the acceptance
region $\mathcal{S}_i$ by constructing a bounding box
$\mathcal{\hat{S}}_i$ around it. The bounding must the part of the
acceptance region $\mathcal{S}_i$ \emph{around} $\thetab_i^*$;
formally $\{ \thetab : d_i(\thetab) < \epsilon$,
$d(\thetab, \thetab_i^*) < M \}$. The condition
$d(\thetab, \thetab_i^*) < M$ describes that if
$\mathcal{S}_i := \{ \thetab : d_i(\thetab) < \epsilon \} $ contains
many \emph{disjoint} sets of $\thetab$ that they fulfill
$d_i(\thetab) < \epsilon$, we want the bounding box to match only the
set that contains $\thetab_i^*$. Ideally, we want the bounding box to
be as tight as possible to $\mathcal{S}_i$, to ensure high acceptance
rate in the importance sampling, and big enough for not discarding
valid parts of $\mathcal{S}_i$.

The bounding boxes are built in two steps; firstly, we define their
axes $\mathbf{v}_d$ in the $d$-dimensional space and, then, the limits
are obtained with a line-search method looking for the smallest
$\kappa: d_i(\thetab_i^* + \kappa \vb_d) \geq
\epsilon$\footnote{$-\kappa$ is used as well for the opposite
  direction along the search
  line}. Algorithm~\ref{alg:region_construction} describes the method
in-depth. After limits are found along all directions, the uniform
distribution $q_i$ is defined on the bounding box, and is used as the
proposal region for importance sampling.

@@

After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
are no longer interested in $d_i(\thetab)$ outside
$\mathcal{\hat{S}}_i$. In the inference steps
(i.e. sampling~\eqref{eq:sampling}, computing
expectation~\eqref{eq:expectation}, evaluating the
posterior~\eqref{eq:approx_posterior_omc}), we need to evaluate the
distance function $d_i$ for checking if the samples obtained from the
uniform distribution $\thetab_{ij} \sim q_i$ lie inside the acceptance
region $\mathcal{S}_i$. Evaluating $d_i$ involves running the
deterministic simulator $M_d(\thetab_{ij}, \vb_i)$, which can be
inneficient when the model is big.\footnote{If Bayesian Optimization
  has been used for obtaining $\thetab_i^*$, then $\hat{d}_i$ can be
  used as an alternative.}. Hence, a \emph{light} surrogate model
$\Tilde{d}_i$ can be fit for approximating $d_i$ around $\theta_i^*$.

ROMC proposes fitting a simple quadratic model\footnote{Any ML
  regression model may be chosen as local surrogate. The choice should
  consider the properties of the local region (i.e. size,
  smoothness). } by sampling training points from the bounding box and
labeling them by evaluating $d_i$. This additional step adds an extra
effort in the training part, but promises a significant speed-up at
the inference phase. Such time-badget imbalance between the training
and the inference part is demanded in many ML contexts; it is usual to
be quite generous with the execution time at the training phase but
quite eager at the inference phase. Fitting a local surrogate model
aligns with this requirement.

### 2021-08-25 latex_files/ch2/2_5_ROMC_algorithmic.tex
minor improvements
changed notation to be inline with ROMC paper

For designing an extendable implementation of ROMC, we firstly define
ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
serves as the driver for the implementation at \pkg{ELFI} that
follows. At a high level, ROMC can be split into the training and the
inference part. In general terms, the training part covers all the
appropriate steps for estimating the proposal regions and the
inference part for getting weighted
samples. Algorithm~\ref{alg:romc_algorithm} defines formally ROMC;
Steps 2-10 (before the vertical line) form the training part, while
steps 12 and 13 the inference part.

\begin{algorithm}[!ht]
	\caption{ROMC}\label{alg:romc_algorithm}
	\begin{algorithmic}[1]
    \Procedure{ROMC}{$M_r(\thetab)$}
    \For{$i \gets 1 \textrm{ to } n_1$}
      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define deterministic distance function}
      \State Obtain $\thetab_i^*, d_i^*$ \Comment{Solve optimization problem}
    \EndFor
    \State $\text{L} \gets \{d_{i, i = 1, ..., n_1}: d_i^* < \epsilon \} $ \Comment{List with filtered solutions}
    \For{$i \gets 1 \textrm{ to } len(L)$}
      \State Construct $q_i$ \Comment{Compute proposal area}
      \State Compute $\Tilde{d}_i$ \Comment{Fit surrogate model in $q_i$}
      \EndFor
    \\\hrulefill
    \State Draw $\{w_{ij}, \thetab_{ij}\}$ \Comment{Draw $n_2$ samples per $q_i$}
    \State Draw $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}    
    \EndProcedure
	\end{algorithmic}
\end{algorithm}

\subsubsection*{Training part}
\noindent
At the training (fitting) part, the goal is the estimation of the
proposal regions $q_i$. The tasks are; (a) sampling the nuisance
variables $\vb_i \sim p(\vb)$ for obtaining the deterministic
functions $g_i(\thetab)$, (b) defining the optimisation problems
$\min_{\thetab} \: g_i(\thetab)$, (c) obtaining $\thetab_i^*$ and
$d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and (e) building
the bounding box for obtaining the proposal region $q_i$. If gradients
are available, using a gradient-based method is advised for obtaining
$\thetab_i^*$ much faster. In case $\nabla_{\thetab} g_i$ is available
in closed-form, the procedure is upgraded in terms of both accuracy
and efficiency; If closed-form description is not available, gradients
approximation with finite-differences requires two evaluations of
$g_i$ for \emph{each} parameter $\theta_d$, which can only work in
low-dimensional problems. If $g_i(\thetab)$ is not differentiable, the
Bayesian Optimisation can be used as an alternative solution. In this
scenario, the training part becomes slower due to fitting of the
surrogate model and the blind optimisation
steps. Algorithm~\ref{alg:training_GB} defines formally the steps of a
gradient-based optimiser and algorithm~\ref{alg:training_GP} the steps
of the Bayesian Optimisation. After obtaning the optimal points
$\thetab^*$, the training part gets completed by estimating the
proposal regions. Algorithm~\ref{alg:region_construction} describes
the line search approach for finding the region's boundaries.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$ or (c) evaluating the
unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Computing an
expectation can be done easily after weighted samples are obtained
using the equation~\ref{eq:expectation}, so we do not discuss it
separately.

\noindent
Sampling is performed by getting $n_2$ samples from each proposal
distribution $q_i$. For each sample $\thetab_{ij}$, the indicator
function is evaluated $\indicator{\regioni(\data)}(\thetab_{ij})$ for
checking if it lies inside the acceptance region. If this is the case,
the corresponding weight is computed as in \eqref{eq:sampling}. As
before, if a surrogate model $\hat{d}$ is available, it can be
utilised for evaluating the indicator function.

\noindent
Evaluating the unnormalised posterior requires the deterministic
functions $g_i$ and the prior distribution $p(\thetab)$\footnote{There
  is no need for solving the optimisation problems and building the
  proposal regions.}. The evaluation requires iterating over all $g_i$
and evaluating the distance from the observed data.

\begin{minipage}{0.46\textwidth}
\begin{algorithm}[H]
    \centering
    \caption{Training Part - Gradient-based. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GB}
    \begin{algorithmic}[1]
      \For{$i \gets 1 \textrm{ to } n$}
        \State Obtain $\thetab_i^*$ using a Gradient Optimiser
        \If{$g_i(\thetab_i^*) > \epsilon$}
        \State{go to} 1
        \Else
        \State Approximate $\jac_i^* = \nabla g_i(\theta)$ and $H_i \approx \jac^T_i\jac_i$
        \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
        \EndIf      
      \EndFor
      \Return{$q_i, p(\theta), g_i(\theta)$}
    \end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{0.46\textwidth}
\begin{algorithm}[H]
    \centering
    \caption{Training Part - Bayesian optimisation. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GP}
    \begin{algorithmic}[1]
      \For{$i \gets 1 \textrm{ to } n$}
        \State Obtain $\thetab_i^*, \hat{d}_i(\thetab)$ using a GP approach
        \If{$g_i(\thetab_i^*) > \epsilon$}
        \State{go to} 1
        \Else
        \State Approximate $H_i \approx \jac^T_i \jac_i$
        \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
        \EndIf      
      \EndFor
      \Return{$q_i, p(\theta), \hat{d}_i(\theta)$}
    \end{algorithmic}
\end{algorithm}
\end{minipage}

\begin{algorithm}[!ht]
	\caption{Computation of the proposal distribution $q_i$; Needs, a model of distance $d$, optimal point $\thetab_i^*$, number of refinements $K$, step size $\eta$ and curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
	\begin{algorithmic}[1]
	\State Compute eigenvectors $\vb_{d}$ of $\hess_i$ {\scriptsize ($d = 1,\ldots,||\thetab ||)$}
	\For{$d \gets 1 \textrm{ to } ||\thetab||$}
		\State $\Tilde{\thetab} \gets \thetab_i^*$ \label{algstep:box_constr_start}
		\State $k \gets 0$
		\Repeat
        	\Repeat
                \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}$ \Comment{Large step size $\eta$.}
        	\Until{$d(f_i(\Tilde{\thetab}), \data) > \epsilon$}
        	\State $\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}$
        	\State $\eta \gets \eta/2$ \Comment{More accurate region boundary}
        	\State $k \gets k + 1$
    	\Until $k = K$
    	\State Set final $\Tilde{\thetab}$ as region end point. \label{algstep:box_constr_end}
    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for $\mathbf{v}_{d} = - \mathbf{v}_{d}$
	\EndFor
	\State Fit a rectangular box around the region end points and define $q_i$ as uniform distribution
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
    \centering
    \caption{Sampling. Requires a function of distance $(g_i(\theta)$ or $\hat{d}_i$ or $\hat{g}_i), p(\theta), q_i$}\label{alg:sampling_GB}
    \begin{algorithmic}[1]
      \For {$i \gets 1 \textrm{ to } n_1$}
      \For {$j \gets 1 \textrm{ to } n_2$}
          \State $\thetab_{ij} \sim q_i$
          \If {$g_i(\thetab_{ij}) > \epsilon$}
            \State Reject $\theta_{ij}$
          \Else {}
            \State $w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}$
            \State Accept $\thetab_{ij}$, with weight $w_{ij}$
          \EndIf
      \EndFor
      \EndFor
    \end{algorithmic}
\end{algorithm}

>>>>

For designing an extendable implementation of ROMC, we firstly define
ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
serves as the driver for the implementation at \pkg{ELFI}. At a high
level, ROMC can be split into the training and the inference part. In
general terms, the training part covers all steps for estimating the
proposal regions and the inference part for getting weighted
samples. Algorithm~\ref{alg:romc_algorithm} defines ROMC formally;
Steps 2-10 (before the vertical line) form the training part, while
steps 12-17 the inference part.

\begin{algorithm}[!ht]
	\caption{ROMC}\label{alg:romc_algorithm}
	\begin{algorithmic}[1]
    \Procedure{ROMC}{$M_r(\thetab), n_1, n_2, \epsilon$}
    \For{$i \gets 1 \textrm{ to } n_1$}
      \State $\vb_i \sim p(\vb)$ \Comment{Draw nuisance variables}
      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define distance function}
      \State $\thetab_i^* = \text{argmin}_{\thetab} d_i$, $d_i^*=d_i(\thetab_i^*)$ \Comment{Solve optimization problem}
      \If{$d^*_i > \epsilon$}
        \State Go to 2 \Comment{Filter solution}
      \EndIf
      \State Define $q_i$ over $\mathcal{\hat{S}}$ \Comment{Estimate proposal area}
      \State (Optional) Fit $\Tilde{d}_i$ on $\mathcal{\hat{S}}$ \Comment{Fit surrogate model}
      \\\hrulefill
      \For{$j \gets 1 \textrm{ to } n_2$}      
        \State $\thetab_{ij} \sim q_i$, compute $w_{ij}$ \Comment{Sample}
      \EndFor
    \EndFor
    \State $E_{p(\thetab|\data)}[h(\thetab)]$ \Comment{Estimate an expectation}
    \State $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}
    \EndProcedure
	\end{algorithmic}
\end{algorithm}

\subsubsection*{Training part}
\noindent
At the training (fitting) part, the goal is the estimation of the
proposal regions $\mathcal{\hat{S}}_i$. The tasks are; (a) sampling
the nuisance variables $\vb_i \sim p(\vb)$ for obtaining the
deterministic functions $d_i(\thetab)$, (b) defining the optimisation
problems $\min_{\thetab} \: g_i(\thetab)$, (c) obtaining
$\thetab_i^*$, $d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and
(e) computing the proposal distribution $q_i$. If gradients are
available, using a gradient-based method is advised for obtaining
$\thetab_i^*$ much faster. In case $\nabla_{\thetab} g_i$ is available
in closed-form, the procedure is upgraded in terms of both accuracy
and efficiency; If closed-form description is not available, gradients
approximation with finite-differences requires two evaluations of
$g_i$ for \emph{each} parameter $\theta_d$, which can only work in
low-dimensional problems. If $g_i(\thetab)$ is not differentiable, the
Bayesian Optimisation can be used as an alternative solution. In this
scenario, the training part becomes slower due to fitting of the
surrogate model and the blind optimisation steps. After obtaning the
optimal points $\thetab^*$, the training part gets completed by
estimating the proposal regions;
Algorithm~\ref{alg:region_construction} describes the line search
approach for finding the region's boundaries.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$ or (c) evaluating the
unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Sampling is
performed by getting $n_2$ samples from each proposal distribution
$q_i$. For each sample $\thetab_{ij}$, the distance
function\footnote{As before, if a surrogate model $\hat{d}$ is
  available, it can be utilised as the distance function.} is
evaluated for checking if it lies inside the acceptance
region. Algorithm~\ref{alg:sampling_GB} defines the steps for
computing a weighted sample. Computing an expectation can be done
easily after weighted samples are obtained using the
equation~\ref{eq:expectation}, so we do not discuss it
separately. Evaluating the unnormalised posterior requires the
deterministic functions $g_i$ and the prior distribution
$p(\thetab)$\footnote{There is no need for solving the optimisation
  problems and building the proposal regions.}. The evaluation
requires iterating over all $g_i$ and evaluating the distance from the
observed data.

% \begin{minipage}{0.46\textwidth}
% \begin{algorithm}[H]
%     \centering
%     \caption{Training Part - Gradient-based. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GB}
%     \begin{algorithmic}[1]
%       \For{$i \gets 1 \textrm{ to } n$}
%         \State Obtain $\thetab_i^*$ using a Gradient Optimiser
%         \If{$g_i(\thetab_i^*) > \epsilon$}
%         \State{go to} 1
%         \Else
%         \State Approximate $\jac_i^* = \nabla g_i(\theta)$ and $H_i \approx \jac^T_i\jac_i$
%         \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
%         \EndIf      
%       \EndFor
%       \Return{$q_i, p(\theta), g_i(\theta)$}
%     \end{algorithmic}
% \end{algorithm}
% \end{minipage}
% \hfill
% \begin{minipage}{0.46\textwidth}
% \begin{algorithm}[H]
%     \centering
%     \caption{Training Part - Bayesian optimisation. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GP}
%     \begin{algorithmic}[1]
%       \For{$i \gets 1 \textrm{ to } n$}
%         \State Obtain $\thetab_i^*, \hat{d}_i(\thetab)$ using a GP approach
%         \If{$g_i(\thetab_i^*) > \epsilon$}
%         \State{go to} 1
%         \Else
%         \State Approximate $H_i \approx \jac^T_i \jac_i$
%         \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
%         \EndIf      
%       \EndFor
%       \Return{$q_i, p(\theta), \hat{d}_i(\theta)$}
%     \end{algorithmic}
% \end{algorithm}
% \end{minipage}

\begin{algorithm}[!ht]
	\caption{Approximation of $\mathcal{\hat{S}}_i$ with a bounding box; Requires, a model of distance $d_i(\thetab)$, optimal point $\thetab_i^*$, number of refinements $K$, step size $\eta\_\text{start}$, maximum iterations $M$ and curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
	\begin{algorithmic}[1]
	\State Compute eigenvectors $\vb_{d}$ of $\hess_i$ {\scriptsize ($d = 1,\ldots,||\thetab ||)$}
	\For{$d \gets 1 \textrm{ to } ||\thetab||$}
		\State $\Tilde{\thetab} \gets \thetab_i^*$ \label{algstep:box_constr_start}
		\State $k \gets 0$
		\State $\eta \gets \eta\_\text{start}$ \Comment{Initialize $\eta$}
		\Repeat
          \State $j \gets 0$
        	\Repeat
            \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}$ \Comment{Large step size $\eta$.}
            \State $j \gets j + 1$
        	\Until{$d(f_i(\Tilde{\thetab}), \data) > \epsilon$ or $j \geq M$} \Comment{Check distance or maximum iterations}
        	\State $\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}$
        	\State $\eta \gets \eta/2$ \Comment{More accurate region boundary}
        	\State $k \gets k + 1$
      \Until $k = K$
      \If{$\Tilde{\thetab} = \thetab_i^*$} \Comment{Check if no step has been done}
        \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d$ \Comment{Then, make the minimum step}
      \EndIf
    	\State Set final $\Tilde{\thetab}$ as region end point. \label{algstep:box_constr_end}
    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for $\mathbf{v}_{d} = - \mathbf{v}_{d}$
	\EndFor
	\State Fit a rectangular box around the region end points and define $q_i$ as uniform distribution
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
    \centering
    \caption{Sampling. Requires a function of distance $d_i$, the prior distribution $p(\thetab)$, the proposal distribution $q_i$}\label{alg:sampling_GB}
    \begin{algorithmic}[1]
      \State $\thetab_{ij} \sim q_i$
          \If {$d_i(\thetab_{ij}) > \epsilon$}
            \State Go to 2 \Comment{Reject sample}
          \Else {}
            \State $w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}$ \Comment{Compute weight}
            \State Store $(w_{ij}, \thetab_{ij})$ \Comment{Store weighted sample}
          \EndIf
    \end{algorithmic}
\end{algorithm}

### 2021-08-25 latex_files/ch3/3_2_1_general_design.tex
minor improvements
changed notation to be inline with ROMC paper

The overview of our implementation is illustrated in
figure~\ref{fig:romc_overview}; one may interpret
figure~\ref{fig:romc_overview} as a depiction of the main class and
the ellipses are the main routines of the class. Following
\proglang{Python}'s naming principles, the methods starting with an
underscore (green ellipses) represent internal (private) functions,
whereas the rest (blue ellipses) are the public methods that the user
interacts with.

In figure~\ref{fig:romc_overview}, it can be easily observed that the
implementation follows consistently the algorithmic description
described above, which follows the partition into training, inference
and evaluation. The training part includes all the steps until the
computation of the proposal regions i.e. sampling the nuisance
variables, defining the optimization problems, solving them,
constructing the regions and fitting local surrogate models. The
inference part comprises of evaluating the unnormalized posterior (and
the normalized when is possible), sampling and computing an
expectation. From an algorithmic point of view, the training and the
inference parts are \emph{isolated}. We also provide some utilities
for inspecting the training process, such as plotting the histogram of
the final distances or visualizing the constructed bounding
boxes. Finally, in the evaluation part, we provide two methods for
evaluating the inference; (a) computing the Effective Sample Size
(ESS) of the samples and (b) measuring the divergence between the
approximate posterior the ground-truth, if the latter is
available.\footnote{Normally, the ground-truth posterior is not
  available; However, this functionality is useful in cases where the
  posterior can be computed numerically or with an alternative method
  (i.e.\ ABC Rejection Sampling), and we would like to measure the
  discrepancy between the two approximations.}

>>>>

The overview of our implementation is illustrated in
figure~\ref{fig:romc_overview}; one may interpret
figure~\ref{fig:romc_overview} as a depiction of the main class and
the ellipses are the main routines of the class. Following
\proglang{Python}'s naming principles, the methods starting with an
underscore (green ellipses) represent internal (private) functions,
whereas the rest (blue ellipses) are the public methods that the user
interacts with. In figure~\ref{fig:romc_overview}, it can be easily
observed that the implementation follows consistently the algorithmic
view presented in~\ref{alg:romc_algorithm}. The training part includes
all the steps until the computation of the proposal regions
i.e. sampling the nuisance variables, defining the optimization
problems, solving them, constructing the regions and fitting local
surrogate models. The inference part comprises of evaluating the
unnormalized posterior (and the normalized when is possible), sampling
and computing an expectation. We also provide some utilities for
inspecting the training process, such as plotting the histogram of the
final distances or visualizing the constructed bounding
boxes. Finally, in the evaluation part, we provide two methods for
evaluating the inference; (a) computing the Effective Sample Size
(ESS) of the samples and (b) measuring the divergence between the
approximate posterior the ground-truth, if the latter is
available.\footnote{Normally, the ground-truth posterior is not
  available; However, this functionality is useful in cases where the
  posterior can be computed numerically or with an alternative method
  (i.e.\ ABC Rejection Sampling), and we would like to measure the
  discrepancy between the two approximations.}
